### 2024-03-09 replies_reviews.txt
sync

----------------------- REVIEW 1 ---------------------

This paper introduces a robustification of a Bayesian approach to indentify 
whether a treatment has an influence on an outcome variable, including possible 
effects of external variables. The model is assumed to be linear (both between 
outcomes and treatments/variables, and between treatment and variables).

Overall, the paper is well written and I believe rigourous (my knowledge of this
particular approach is limited). I would say the technical extension is rather 
simple (considering sets of parameters and performing sampling experiments wihtin 
these sets, followed by a classical Bayesian analysis), and the experiments 
synthetic, hence the most interesting part of the paper appears to me as the one 
trying to identify why going imprecise/robust is good. However the analysis of 
the experiments is a bit cryptic to me. Still, the paper seems ok enough to me to 
be presented and discussed at ECSQARU, and it fits the scope of the conference.

-> Thank you very much for your positive feedbacks. We have modified the 
-> simulation studies to explain the methods in a better way.

Given my limited knowledge of the formal model, most of my relevant comments 
will concern the experiments:

* I do not really understand what is reported in Table 2: what is the quantity 
we read, what should be his ideal value (lower=better, higher=better, fitting a
particular estimate?). Form my understanding, this is the estimate of \beta_T 
once the Bayesian analysis has been achieved. This should be clearly stated. 
The results however seem to show that the Bayesian analysis is doing really fine 
in both cases, and it would also be interesting to report some credibility 
intervals for those. It would also probably be interesting to look what happens 
in case of model/prior misspescification, and how the two approaches would react 
in this case.

-> We propose a robust Bayesian analysis to show its benefits in situation
-> where we have limited data to perform our analysis. We now have modified
-> the text and the table to make it more interpretable.

* Table 3 is a bit frustrating, as 1. we also appear to be in a rebgime where 
classical approaches mostly do fine (maybe finding a regime where imprecision 
works better would be useful?), 2. it is not clear what are the origins/main
contributors to the loss when it is non-null (mostly false positives/false 
negatives or abstention?) and 3. what would give the classical Bayesian approaches 
if one used the proposed loss functions to take decisions, using a rather classical
appraoch of seeing abstention as a peculiar decision in competition with the 
others... albeit by choosing \ell_3=0.5 which is similar to discounted accuracy 
it is quite likely that such an otpion would never be considered by precise methods.

-> Thank you very much for this interesting feedback. As we mentioned earlier,
-> we are particularly focusing on the situations where we have limited data.
-> Since it is a regressional framework, the analyses tends to support
-> classical approaches for $n \ge 50$. We have modified the simulation
-> setting to show results for $25 \le n \le 75 $, to reflect a regime where
-> imprecision is more useful.

Minor comments:

* Equations (2) and (3): I am not sure what to do subject-specific random variable? 
How can we interpret a single individual experiment as a random quantity (I know 
this is classical in statistics and that "in theory" everything goes fine, but in
practice this seems weird).

-> The random variables for subject specific
-> quantities, such as Yi and Ti, denote quantities before they are observed.
-> Since these quantities, before observation, are uncertain, they must be treated
-> as random variables. For instance, the decision to treat or not, Ti, could be
-> based on some bio markers, which are inherently random due to noise.
-> The uncertainty about the response Yi could be more epistemic in nature,
-> although there could also be random factors that influence the response.
-> The resulting subject specific values of delta_i correspond then to a
-> mathematical formulation of treatment effect heterogeneity.

* I am not sure Table 1 information should be a table rather than in-line texts. 
It seems to me like its taking space without adding much structure/info.

-> We now have modified the table accordingly.

----------------------- REVIEW 2 ---------------------

The authors focus on causal inference, and the goal here is to enlarge the usual
Bayesian analysis so that a set of probability distributions is contemplated and 
so that robustness can be assessed. This is a relevant issue (even though it is
presented as if its relevant is obvious, and some words connecting with the bursting
literature on causal thinking in AI would be really valuable). And the paper is 
well written and easy to follow. The contribution is somewhat narrow as it basically
consists in taking an existing model and adding a set of probabilities to it, so 
as to examine robustness and sensitivity. This makes sense but the idea seems quite
natural and does not seem to bring technical insight (perhaps this should be 
discussed: is there any technical issue that is challenging in doing the extension
proposed in the paper?). The weak part is the validation, as it is based on apparently
simple simulations. It would be a lot more convincing to use real data; given the
amount of data available today, it would be strange not to find some dataset that 
is appropriate. Any thoughts?

-> Thank you for the comment. We agree it would be better to use real data.
-> Unfortunately we cannot fully implement this suggestion for the conference paper,
-> but we have acknowledged this issue in the conclusion, and we shall definitely take
-> your suggestion forward in future work.

Concerning the text, a few suggestions:
- The abstract is rather long and wordy; perhaps it could be shortened and made sharper.

-> We agree. We have made the abstract a bit shorter.

- A few places: "can not" instead of "cannot".
- Expressions (20) and (21) should have ending punctuation.
- Page 9: "under estimates" instead of "underestimates"; also, expression "the near
vacuous of set of" seems incorrect.
- References:  "Bayesian" instead of "bayesian".

-> Thank you very much for pointing out the typos. We have corrected those accordingly.

----------------------- REVIEW 3 ---------------------

This is paper focuses on Bayesian hierarchical methods for the estimation of causal effects. The authors adopt a standard regression model for causal estimation and a Bayesian hierarchical model based on that model. The goal is to cope with data scarcity, and a robust approach consisting of a sensitivity analysis wrt the priors is used. This gives a procedure for covariate selection.  A utility approach is also sketched to evaluate whether the available data are sufficient for a robust decision. The method is tested on a synthetic model, and the results compete with existing approaches.

I understand the work of the authors as an application of ideas from robust Bayes and imprecise probability theories to causal analysis. This looks very promising to me. Even if I regard the present work as a first attempt in this direction, the paper might trigger interesting discussions at the conference.

-> Thank you for your positive comments.

I miss some discussions about the optimisation method. I understand everything here is based on sampling, but the differences between standard approaches should also be emphasised. I also don't understand whether the bound approximation can be characterised (e.g., inner approx).

-> Thank you for the comment. We agree it would be better to investigate further theoretical
-> properties and obtain approximations of the posterior estimates.
-> Unfortunately we cannot finish the implementation for the conference paper,
-> but we have acknowledged this issue in the conclusion, and we shall definitely take
-> your suggestion forward in future work.

More generally speaking, the authors might also be interested in the works about the bounding of counterfactuals. Sampling methods have been proposed for that (see various recent works by Bareinboim and colleagues and also Zaffalon and colleagues).

-> Thank you for this interesting insight, currently this is beyond the scope of our
-> contribution but we will definitely take your suggestion for the extension of the
-> paper.