### 2024-03-09 ecsqaru.tex

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[oribibl]{llncs}
%
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{bm}
\usepackage{amsfonts}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{mathtools}
\usepackage{blindtext}
\usepackage[capitalise]{cleveref}
\usepackage[british]{babel}
\usepackage{geometry}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{longtable}
\usepackage{bigints}
%\usepackage{siunitx}
\usepackage{soul}
\usepackage{tikz}
\usepackage{color}
\usepackage{easyReview} % for \comment, \alert, etc.
\usepackage{multicol}
\usepackage{multirow}
\newcommand{\x}{\boldsymbol{X}}
\renewcommand{\b}{\hat{\beta}}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\normal}{\mathcal{N}}
\newcommand{\lexp}{\underline{\text{E}}}
\newcommand{\uexp}{\overline{\text{E}}}

\begin{document}

\title{A Robust Bayesian Approach for Causal Inference Problems
%\thanks{}
}
%
\author{Tathagata Basu\inst{1}\orcidID{0000-0002-6851-154X} \and
Matthias C.~M.~Troffaes\inst{2}\orcidID{0000-0002-1294-600X} \and
Jochen Einbeck\inst{2,3}\orcidID{0000-0002-9457-2020}}
%
\authorrunning{Basu et al.}
%
\institute{Civil and Environmental Engineering, University of Strathclyde, UK \and
Department of Mathematical Sciences, Durham University, UK \and
Durham Research Methods Centre, UK}
%
\maketitle              
%
\begin{abstract}
Causal inference
%using observational data is an important problem in
%many fields such as epidemiology, social science, economics, etc.
%Our goal is to find
concerns finding
the treatment effect on subjects
along with causal links between the variables and the outcome.
However,
the underlying heterogeneity between subjects makes the
problem practically unsolvable. Additionally, we often
need to find a subset of
explanatory variables to understand the treatment effect.
Currently, variable selection methods tend to maximise the
predictive performance of the underlying model, and unfortunately,
under limited data, the predictive performance is hard to assess,
leading to
harmful consequences. To address these issues, in this paper, we consider a
robust Bayesian analysis which accounts for abstention in selecting
explanatory variables in the high dimensional regression model.
To achieve that, we consider a set of spike and slab priors
through prior elicitation to obtain a set of posteriors for
both the treatment and outcome model. We are specifically interested
in the sensitivity of the treatment effect in high dimensional causal
inference as well as identifying confounder variables.
However, confounder selection can be deceptive
in this setting, especially when a predictor is strongly associated
with either the treatment or the outcome.
To avoid that we apply a post-hoc selection scheme,
attaining a smaller set of confounders as well as separate sets of variables
which are only related to treatment or outcome model. Finally, we illustrate
our method to show its applicability.

\keywords{high dimensional data \and variable selection \and Bayesian analysis \and imprecise probability.}
\end{abstract}

\section{Introduction}\label{sec:intro}

In causal inference, we are interested in estimating the causal
effect of independent variables on a dependent variable. Ideally,
randomised trials are the most efficient way to perform this task.
However, this is not always practical for several reasons; ethical 
concerns, design cost, population size, to name a few. This
leaves us with observational studies which are usually obtained
by means of collecting data though surveys or record keeping. But this
can be problematic in the presence of confounders, which are variables associated with both the treatment and the outcome.
In such cases, we need to be extra cautious as otherwise it will
lead to unwanted bias in the treatment effect estimator \cite{rosenbaum83}.
Several works have been done in order to tackle the presence of
confounder variables. One such work in the topic was 
by Robins \cite{Robins1986ANA} where the author used  a graphical
approach for the identification of the causal parameters.
Rosenbaum and Robin \cite{rosenbaum1985} suggested the use of a link model to estimate
the propensity scores for all individuals. Later on several other
methods have been proposed based on propensity score matching.
A brief review on such methods can be found in \cite{winship99, stuart10}.

The Bayesian approach in causal effect estimation is a popular
strategy in the field and one of the earlier works on this can be
found in \cite{rubin1978}. Lately, with the rise of high dimensional 
data, Bayesian methodologies have become
more appealing. Crainiceanu \text{et al.~}\cite{Crainiceanu2008} proposed a bi-level 
Bayesian model averaging based method for estimating the causal 
effect. Wang \text{et al.~}\cite{wang2015} suggested BAC (or, Bayesian adjustment for
confounding) where they use an informative prior obtained from
the treatment model and apply them on the outcome model for
estimating causal effect. Several other methods were also
proposed to tackle confounders from the point of view of Bayesian
variable selection, see for instance \cite{Zigler2014,Hahn2018} among others.

In this paper we take inspiration from the approach of Koch \text{et al.~}\cite{koch2020}, who proposed a bi-level spike and slab prior for causal effect 
estimation. They considered a data-driven adaptive approach to
propose their prior which reduces the variance of the causal estimate. 
In our approach, we perform a 
sensitivity analysis based approach where instead of using a single prior, 
we consider a set of priors \cite{BERGER1990303}. This is particularly 
interesting as in many cases, causal effect estimation can be performed 
through a meta analysis and hence robust Bayesian analysis \cite{raices_cruz22} 
can be beneficial under severe uncertainty. Moreover, for some problems 
we have to rely on very limited data to perform our Bayesian analysis and 
inference may not be reliable in presence of heteroscedasticity within
the data. Instead, we use expert opinion and elicit a set of priors 
based on empirical evidence. 
This also allows us to construct the problem of confounder identification 
in a framework where abstention has a relatively positive gain i.e.~
when the cost of further tests/data collection is cheaper than
mistreating a subject. To propose our 
framework, we consider a set of continuous spike and slab priors 
\cite{ishwaran2005} for confounder identification and construct a Bayesian 
group LASSO \cite{xu2015} type problem. To perform the prior sensitivity 
analysis, we consider a set of beta priors on the covariate selection 
probability of the spike and slab priors. We use the posteriors of this
covariate selection probability for identifying the confounders. Finally, 
we consider a post-hoc coefficient adjustment method \cite{hahn2015}
to recover sparse estimates associated with either the outcome or the
treatment model.

The rest of the paper is organised as follows. In \cref{sec:causal}
we give a formal description of the causal estimation problem in the
context of linear regression. \Cref{sec:bayes} is focused on the
Bayesian analysis of causal inference problems, followed by the
motivation of a robust Bayesian analysis along with our proposed decision 
theoretic framework for confounder (variable) selection. In \cref{sec:sim}, 
we provide results of simulation studies under different scenarios 
and show the possible applications in real life problems. Finally, 
we discuss our findings and conclude this paper in \cref{sec:conc}.

\section{Causal Estimation}\label{sec:causal}

Let an observational study give us the outcomes $Y=(Y_1, \dots, Y_n)$ along with 
corresponding treatment indicators $T=(T_1, \dots, T_n)$. Then the
treatment 
effect in the population is given by the expectation of the difference
in outcomes between the treatment and controls:
\begin{align}
	\delta = \mathbb{E}(Y\mid T =1) - \mathbb{E}(Y\mid T=0).
\end{align}
Similarly, the individual causal
effect of treatment $T_i$ on outcome $Y_i$ is given by:
\begin{align}
	\delta_i \coloneqq \mathbb{E}(Y_i\mid T_i=1) - \mathbb{E}(Y_i\mid T_i=0).
\end{align}
That is, we are interested in the difference between the outcomes when the $i$-th subject receives
the treatment and when 
%$i$-th subject 
it remains as a control.

In theory, both of these quantities exist.
However, we cannot observe $\mathbb{E}(Y_i\mid T_i=1)$ and $\mathbb{E}(Y_i\mid T_i=0)$
the average causal effect of the treatment $T$ by calculating the averaged
outcome of all the subjects that received the treatment and
all the subjects that remained as control:
\begin{align}
	\hat{\delta} \coloneqq 
	\frac{\sum_{i=1}^n Y_i\cdot\mathbb{I}(T_i=1) - 
		\sum_{i=1}^n Y_i\cdot\mathbb{I}(T_i=0)}{n}.
\end{align}
However, this relies on an important assumption that the treatment effect
on the $i$-th subject given that they received the treatment is the same as the (counterfactual) treatment effect when they remain as control \cite{winship99}.

\subsection{Regression Model}
Regression methods are widely used in causal effect estimation. The
main idea behind these regression methods is to remove the
correlation between the treatment indicator and the error term
\cite{winship99,HECKMAN1985}. To do so, we rely on $p$ different observed quantities
or predictors denoted by $X\coloneqq$ $[X_1^T, \dots, X_n^T]^T$
where each $X_i\in\mathbb{R}^p$.
Each $X_i$ is treated as a $p$-dimensional row vector,
so $X$ is a $n\times p$ matrix.
Now, let
$\beta \coloneqq (\beta_1$, \dots, $\beta_p)^T$ denote the vector of regression
coefficients
related to the predictors, and $\beta_T$ denote a regression coefficient related to the treatment.
Then we can define a linear model for the outcome
so that
\begin{equation}
	Y_i =  T_i \beta_{T} + X_i\beta + \epsilon_i
\end{equation}
where $\epsilon_i\sim \mathcal{N}(0, \sigma^2)$. Clearly, when
the underlying true outcome model is linear with respect to the treatment,
\begin{equation}
	\delta_i = \mathbb{E}(Y_i\mid T_i =1) - \mathbb{E}(Y_i\mid T_i=0) = \beta_{T}.
\end{equation}
In the presence of confounders we also need to consider the
association between the treatment indicators and the predictors.
In literature, authors often suggest a probit link function to 
construct the regression model. This way, we can
specify the conditional probability that subject $i$ receives the treatment through a linear model. 
That is, for another vector of regression coefficients 
$\gamma\coloneqq(\gamma_1, \cdots, \gamma_p)^T$ we define
\begin{align}
	P(T_i=1\mid X_i) = \Phi(X_i\gamma)
\end{align}
where $\Phi(\cdot)$ denotes the cumulative distribution function
of a standard normal distribution. To incorporate this probit
link function, we assume that we can model the $T_i$ through the following \cite{albert93}:
\begin{align}
    T_i^* &= X_i\gamma +u_i \\
    T_i   &= \mathbb{I}(T_i^*>0)
    =
    \begin{cases}
    1 & \text{if }T_i^*>0 \\
    0 & \text{otherwise}
    \end{cases}
\end{align}
where $u_i\sim\mathcal{N}(0,1)$.

Now, to construct the joint likelihood function, we define an extended
output $2n\times 1$ column vector
$W\coloneqq\left(\begin{smallmatrix}Y \\ T^*\end{smallmatrix}\right)$
and corresponding $2n\times(2p+1)$ dimensional design matrix
\begin{align}
	Z &\coloneqq
        \begin{bmatrix}
           T_1 & X_1 & 0 \\
           \vdots & \vdots & 0 \\
           T_n & X_n & 0 \\
           0 & 0 & X_1 \\
           \vdots & \vdots & \vdots \\
           0 & 0 & X_n
        \end{bmatrix}
        =
	\begin{bmatrix}
		X_O & 0 \\
		0 & X_T
	\end{bmatrix}
\end{align}
where, $X_O = [T, X]$ and $X_T = X$. Then, considering the assumption of
Gaussian error terms, we have the following likelihood distribution
\begin{align}
	W\mid Z, \beta_T, \beta, \gamma, \sigma^2 \sim\normal\left(Z\nu, \Sigma\right)\label{eq:like:group},
\end{align}
where $\nu = (\beta_T, \beta^T, \gamma^T)^T$ and
\begin{align}
	\Sigma &=
	\begin{bmatrix}
		\sigma^2{I}_n & 0 \\
		0 & {I}_n
	\end{bmatrix}.
\end{align}

\section{Bayesian Causal Estimation}\label{sec:bayes}

The likelihood given by \cref{eq:like:group} gives us
a foundation for a Bayesian group LASSO 
\cite{xu2015} type model. This way, we can look into the posterior selection
probability associated with the $j$-th predictor. There are several
ways to construct spike and slab priors which achieve 
variable selection. In our case, we consider a continuous type
\cite{ishwaran2005} prior for faster posterior
computation.

\subsection{Hierarchical model}

Let $\pi_j$ denote the prior probability that the $j$-th
predictor is associated to the outcome or the 
treatment. That is, 
\begin{equation}
	\pi_j = P\left((\beta_j,\gamma_j)\not=(0,0)\right).
\end{equation}
Then we can define the following hierarchical model for spike and 
slab group LASSO so that,
for $1\le j\le p$,
\begin{align}
	(\beta_j,\gamma_j)^T \mid \pi_{j}, \sigma^2 &\sim 
	\pi_{j}\normal\left( \begin{bmatrix}
		0 \\
		0
	\end{bmatrix}, 
	\tau_1^2\begin{bmatrix}
		\sigma^2 & 0 \\
		0 & 1
	\end{bmatrix}\right)
	+ (1-\pi_{j}) \normal\left(\begin{bmatrix}
		0 \\
		0
	\end{bmatrix}, 
	\tau_0^2\begin{bmatrix}
		\sigma^2 & 0 \\
		0 & 1
	\end{bmatrix}\right)\\
	\beta_T\mid \sigma^2 &\sim \normal\left(0, \sigma^2\right)\\
	\frac{1}{\sigma^2}&\sim \text{Gamma}(a, b)\\
	\pi_{j} &\sim\text{Beta}\left(sq, s(1-q)\right).
\end{align}
In the hierarchical model, we fix sufficiently small $\tau_0$
$(1\gg\tau_0>0)$ so that  $(\beta_j, \gamma_j)$ has its probability mass 
concentrated around zero. Therefore, this represents the spike component of our prior specification. 
For the slab component, we consider $\tau_1$ to be large so that $\tau_1\ge 1$. This allows the prior for $(\beta_j,\gamma_j)$ to be flat, besides the spike component at the origin.. 
We illustrate the components of a bivariate spike and slab prior in 
\cref{fig:ssbl} (with fixed $\sigma=1$). We generate the spike component 
with $\tau_0=0.001$ and the slab component with $\tau_1=5$.

\begin{figure}[h]
	\begin{center}
		\includegraphics[width = 0.95\linewidth]{spike_slab_bi.pdf}
	\end{center}
	\caption{Spike (left) and slab (right) components of a bi-variate distribution for $\tau_0 = 0.001$, $\tau_1$ = 5 and $\sigma=1$.}
	\label{fig:ssbl}
\end{figure}

For the precision term $1/\sigma^2$, a natural choice of prior is the gamma distribution
as it allows the control of both the location and the scale of the precision. To
ensure that the prior is able to represent the data, we consider $b=1$ and 
fix $a$ so that it represents the prior mean of the precision. In cases where we have no prior information,
we can simply consider a large value for $a$ so that the interval
$[0, 2a]$ contains the true value of the precision.
As defined earlier, $\pi_j$ is used as the selection probability
of the $j$-th predictor in either of the models and we use a beta prior to specify 
these selection probabilities where  $q_j$ represents our prior expectation of the selection probability ($\pi_j$) and $s$ acts as 
a concentration parameter.
For the causal effect, we want to use a Gaussian distribution that 
matches the scale of the noise term. Therefore, we consider $\beta_T\sim \normal(0,\sigma^2)$.

In \cref{fig:regress}, we show a probabilistic graphical representation
of our hierarchical model. In the figure, grey circular nodes represent the
prior hyper-parameters which will be used for sensitivity analysis
of the model. The transparent circular nodes are used to denote
the modelling parameters which are our quantities of interest. 
The observed quantities are denoted with transparent rectangular
nodes. We also use a grey rectangular node to denote the intermediate
latent variable $T^*$. We use directed edges to denote the
relationship between different nodes. However, we use a dashed
edge between $X$ and $T$ as they are related through the latent
variable $T^*$.

\begin{figure}
	\centering
	\begin{tikzpicture}[params/.style={circle, draw=black!60, very thick, minimum size=7mm},
		hyper/.style={circle, draw=black!60, fill=black!20, thick, minimum size=7mm},
		post/.style={circle, draw=black!60, fill=green!20, thick, minimum size=7mm},
		latent/.style={rectangle, draw=black!60, fill=black!10, dashed, minimum size=7mm},
		data/.style={rectangle, draw=black!60, thick, minimum size=8mm}]
		\node[params] (1) at (0,0) {$\pi$};
		\node[data] (2) at (3,0) {$X$};
		\node[data] (3) at (6,0) {$T$};
		\node[params] (4) at (1.5,1.5) {$\gamma$};
		\node[latent] (5) at (4.5,1.5) {$T^*$};
		\node[params] (6) at (1.5,-1.5) {$\beta$};
		\node[data] (7) at (3.5,-1.5) {$Y$};
		\node[params] (8) at (0,-3) {$\sigma^2$};
		\node[params] (10) at (6,-1.5) {$\beta_T$};
		\node[hyper] (11) at (-1.5,-.8) {$s$};
		\node[hyper] (12) at (-1.5,.8) {$q$};
		\node[hyper] (13) at (-1.5,-2.2) {$a$};
		\node[hyper] (14) at (-1.5,-3.8) {$b$};
		\draw[black, dashed] (0.75,-3.7) rectangle (7,2.1);

		\path (1) edge[->]  (6);
		\path (1) edge[->]  (4);
		\path (8) edge[->]  (6);
		\path (6) edge[->]  (7);
		\path (2) edge[->]  (7);
		\path (2) edge[->]  (5);
		\path (5) edge[<-] (4);
		\path (5) edge[->] (3);
		\path (3) edge[->]  (7);
		\path (10) edge[->]  (7);
		\path (2) edge[dashed][->] (3);
		\path (8) edge[bend right = 60][->]  (10);
		\path (11) edge[->]  (1);
		\path (12) edge[->]  (1);
		\path (13) edge[->]  (8);
		\path (14) edge[->]  (8);

	\end{tikzpicture}
	\caption{Probabilistic graphical representation for causal inference with Bayesian hierarchical model.}
	\label{fig:regress}
\end{figure}

\subsection{Robust Bayesian Analysis}

The hierarchical model presented above is a standard spike and slab model for
variable selection and performs well when we have sufficient data to begin with. 
However, especially in the case of causal inference having sufficient data may not 
be feasible. Moreover, 
we also need to be cautious about the side effects of a treatment.
Therefore, we are particularly interested in constructing a robust Bayesian framework
for variable selection. This way, when we are preparing a guideline for treatment, we 
can have the option to ask for more data before reaching any conclusion. To achieve this,
we consider a utility based framework with three possible ways of determining a variable.

In general, an unsuccessful treatment of a subject can have severe consequences 
which cannot be associated with a suitable loss function. Instead, we
assume that we can always revert any initial mistreatment by further treatments, 
and we can associate a loss function with the cost of further treatments. This way, 
in the simplest case, we can associate two constant loss values $\ell_1,\ell_2$ 
with false positives and false negatives respectively. 
Clearly, false positives will lead to unwanted side effects and
false negatives will lead to mistreatment of the patient. Finally, we associate
a loss value $\ell_3$ for abstention which can be interpreted as the cost of further tests.
Ideally, in most cases, $\ell_3\ll \ell_1,\ell_2$. However, in certain scenarios,
this might not be the case, especially when the condition of a subject deteriorates rapidly
over time.
\iffalse
\begin{table}[h]
    \caption{Losses with respect to different types of variable selection}
    \label{tab:loss:vs}
    \centering
    \begin{tabular}{l||c|c|c}
       &  False Positive & False Negative & Abstention \\
       \hline
       \hline
       Value  &  $\ell_1$ & $\ell_2$ & $\ell_3$
    \end{tabular}
\end{table}
\fi

Now, based on this notion of abstaining from selecting a variable, we can perform
a sensitivity analysis over a set of priors on the prior selection probability.
That is, we can consider a set of possible values for $q$ such that
$q\in\mathcal{P}$, where $\mathcal{P} \subseteq \left(0, 1\right)^{p}$.
Here, the equality occurs for the near vacuous case. However, in real-life
situations, performing a robust Bayesian analysis for the near vacuous case is 
not practical. Instead, we incorporate expert elicitation to define our model.
For instance, we can consider $q\in \left[\underline{q}, \overline{q}\right]$
where $p\underline{q}$ and $p\overline{q}$ represent the bounds of the prior expectation on the
total number of variables present in either of the models.

\subsection{Variable selection and coefficient adjustment}
For the co-variate selection, we look into the posterior expectation of $\pi_j$. 
We consider the $j$-th predictor to be removed from both the
treatment and outcome model, if
\begin{align}\label{eq:vs:remove}
	\uexp (\pi_j\mid W)\coloneqq \sup_{q\in \mathcal{P}} \mathbb{E}(\pi_j\mid W) < 1/2.
\end{align}
Similarly, we consider the $j$-th predictor to be present in at least one of the models, if
\begin{align}\label{eq:vs:sel}
	\lexp (\pi_j\mid W)\coloneqq \inf_{q\in \mathcal{P}} \mathbb{E}(\pi_j\mid W) \ge 1/2.
\end{align}
Otherwise, we consider the variable to be indeterminate,  in which case we abstain from putting
it in any of the models but instead just report a lack of information.

In general, this framework is self sufficient for variable selection. However, for
model fitting and prediction, we need to evaluate the values 
of the regression coefficients. For that we first need to find the set of active
predictors with respect to our prior expectation of the selection probability $q$.
For any fixed $q$, we define the set $S(q)$ as the set of all variables which are active
in the treatment model or in the outcome model:
\begin{equation}
	S(q)\coloneqq
	\left\{j\colon E(\pi_j\mid W) \ge 1/2\right\}.
\end{equation}
For sensitivity analysis,
the intersection of $S(q)$ over all $q$ gives us the set of
active variables obtained through \cref{eq:vs:sel}.
Similarly, the union gives us the set of
variables that are not removed through \cref{eq:vs:remove}.
That is:
\begin{align}
    \mathcal{S}_*\coloneqq \left\{j:\lexp (\pi_j\mid W)\ge1/2\right\}
    = \bigcap_{q\in \mathcal{P}}S(q),
    \qquad \mathcal{S}^*\coloneqq \left\{j:\uexp (\pi_j\mid W)\ge1/2\right\}
    = \bigcup_{q\in \mathcal{P}}S(q).
\end{align}
Clearly, $\mathcal{S}_*\subseteq\mathcal{S}^*$.
$\mathcal{S}_*$ represents the set of variables that are sure to be selected,
$\{1,\dots,p\}\setminus\mathcal{S}^*$ represents the set of variables that are sure to be removed, and
$\mathcal{S}^*\setminus\mathcal{S}_*$ represents the set of variables about which we are undecided.
In this way, through sensitivity analysis, our approach incorporates robustness.

Now, for each fixed value of $q$, let $\hat{\beta}_{S(q)}$ be the posterior means 
of the regression coefficients of the outcome model with respect to
the predictors that belong to $S(q)$. Similarly,
$\hat{\gamma}_{S(q)}$ be the posterior means of the regression
coefficients for the treatment effects. Since we use continuous 
spike and slab priors, these regression coefficients are not sparse.
Moreover, with our variable selection we only determine whether the variable 
is included in at least one of the models. But, we cannot determine a specific 
association. Therefore, to adjust the sparsity of the estimates and understand
the specific association with the treatment/outcome/both, we apply the 
``decoupled shrinkage and selection'' method proposed by \cite{hahn2015}. 
For that, we solve the following adaptive LASSO-type \cite{Zou2006}
problems

\begin{align}
	\hat{\beta}^D_{S(q)} &= 
	\arg\min_{\beta_{S(q)}} \frac{1}{n}\|X_{S(q)}\hat{\beta}_{S(q)}
	- X_{S(q)} \beta_{S(q)}\|_2^2 + \lambda\sum_{j\in S(q)} 
	\frac{|\beta_{j,S(q)}|}{|\hat{\beta}_{j,S(q)}|}
\end{align}
and
\begin{align}
	\hat{\gamma}^D_{S(q)} &= 
	\arg\min_{\gamma_{S(q)}} \frac{1}{n}\|X_{S(q)}\hat{\gamma}_{S(q)}
	- X_{S(q)} \gamma_{S(q)}\|_2^2 + \lambda\sum_{j\in S(q)} 
	\frac{|\gamma_{j,S(q)}|}{|\hat{\gamma}_{j,S(q)}|}
\end{align}
where $q\in \mathcal{P}$.

\iffalse
The DSS method explained
above give us adjusted coefficient estimates for the treatment
and outcome model. However, as result the causal effect estimate
remains the same and modelling with such adjusted sparse effect 
will contribute to the prediction error. Therefore, we need to 
adjust the causal effect estimate as well. Let
\begin{equation}
    \hat{T}^D = \mathbb{I}\left(X_{S(q)}\hat{\gamma}^D_{S(q)} >0\right)
    \quad\text{and}\quad
    \hat{T} = \mathbb{I}\left(X\hat{\gamma}^q >0\right).
\end{equation}
and let $\beta_{T}^D(q)$
be the causal effect after applying `DSS' and $\hat{\beta}_{T}(q)$ be the
posterior estimate of the causal effect for any fixed $q\in\mathcal{P}$. Then
the adjusted estimate is given by:
\begin{equation}
    \beta_{T}^D(q) = \left(\left(\hat{T}^D\right)^T \hat{T}^D\right)^{-1}
    \left(\hat{T}^D\right)^T \left[\hat{\beta}_{T}(q)\hat{T} 
    + X_{S(q)}\left(\hat{\beta}_{S(q)} - \hat{\beta}^D_{S(q)}\right) 
    +X_{S^C(q)}\hat{\beta}_{S^C(q)}\right]
\end{equation}
where
\begin{equation}
    S^C(q)\coloneqq \left\{1,2,\cdots, p\right\} \neg S(q).
\end{equation}
\fi

\section{Simulation Studies}\label{sec:sim}

For the simulation studies, we consider 2 different settings. In each
case, we generate the design matrix $X$ such that $X_i\sim\mathcal{N}(0, \Sigma)$
for $1\le i\le n$ where $[\Sigma]_{ij} = 0.3^{|i-j|}$. This way, we 
generate 50 predictors for our model with mild correlations among them.
We then use the following generation schemes to generate the outcome and
treatment indicator: 
\begin{equation}
    T_i \sim \text{Bernoulli}\left(1/(1+\exp(-X_i\gamma))\right)
    \quad\text{and}\quad
    Y_i = 4T_i + X_i\beta.
\end{equation}
\begin{description}
    \item[Scenario 1] --- $|\gamma_j|, |\beta_j|>0$ for $j\le 10$
    \item[Scenario 2] --- $|\gamma_j|>0$ for $j\le 10$ and $|\beta_j|>0$ for $j\le 15$
\end{description}
For both
the cases, we consider different numbers of observations $n$ where
$n=25+ 5k$ for $k=0,1,2,\cdots,10$. %depends on finishing
%$n=25\times 2^k$ for $k=0,1,2,\cdots,5$.

We present our analyses in \cref{tab:causal} and \cref{tab:misspec}.
For the sake of clarity we use the following accronyms: RBCE for 
robust Bayesian causal estimation (our method); SSCE for spike and
slab causal estimation \cite{koch2020}; BSSCE for bi-level spike and slab causal
estimation \cite{koch2020}; and BSSL for Bayesian spike and slab lasso
\cite{xu2015}. As it can be
seen from both the tables, SSCE and BSSCE are formulated for problems
where $p\le n$ and therefore we do not have any results for $n<50$.

\paragraph{Elicitation} For the
elicitation of $\mathcal{P}$, we use marginal correlation between 
$Y$ and $X$ to determine the bounds on number of active variables. We 
set the thresholds to be $0.15$ and $0.35$ for the correlations. We compute the
number of variables with marginal correlation greater than 0.15
(say $p_1$) and number of variables with  marginal 
correlation greater than 0.35 (say $p_2$). 
We use these numbers to obtain the bounds on the number of active
variables so that $\mathcal{P}=[p_2/p , p_1/p]$.

\paragraph{Initialisation} 
To implement our method, we use \texttt{rjags} and for the other three
methods we use the code provided in the appendix of \cite{koch2020}.
For our method, we set $\tau_0=10^{-6}$ and $\tau_1=1$ to construct the
spike and slab prior. For the noise term, we set $a=10$ and $b=1$.  To perform 
our Bayesian analysis with \texttt{rjags}, we first consider an adaptive 
stage with 2000 iterations followed by discarding of 2000 burn in samples 
to refine the posteriors. We consider 5000 MCMC samples to compute the
posterior estimates. For the other methods we use the in-built settings 
to initiate the analyses.

\paragraph{Results}
We provide our result for causal estimate in \cref{tab:causal}. 
As we perform a sensitivity analysis,
our method gives an interval estimate for the causal effect and we show
that in two different rows where the first row gives the lower bound
and the second row gives the upper bound. We notice that our method is 
somewhat in agreement with the other methods but much more consistent
in terms of estimating the treatment effect. However, this is not the
case for other methods and sometimes those methods produce extreme 
values. This can be observed in \cref{fig:comp:trt} as well. Here, the true 
value is represented by the straight line for $\beta_T = 4$.

From the figure, we can notice that our method tends
to underestimate the causal effect. This suggests that
we may want to have a different value of $a$ for these sets of observations
instead of a fixed value of $a=10$ for all of our analyses.
We can also see that the lower bound tends to improve
with increasing number of observations which validates the assumption that as we accumulate
more information, the interval becomes smaller and converges towards
the true value.

% latex table generated in R 4.3.0 by xtable 1.8-4 package
% Tue May 23 18:29:20 2023
\begin{table}[ht]
\caption{Causal estimates obtained from different methods
for 6 different
numbers of observations.}\label{tab:causal}
\centering
\subcaption*{First scenario: $|\gamma_j|, |\beta_j|>0$ for $j\le 10$}
\begin{tabular}{l|rrrrrrrrrrr}
  \hline
 & 25 & 30 & 35 & 40 & 45 & 50 & 55 & 60 & 65 & 70 & 75 \\ 
  \hline
RBCE (low) & 3.22 & 3.54 & 3.30 & 3.66 & 3.77 & 3.80 & 3.85 & 3.89 & 3.90 & 3.91 & 3.89 \\ 
  RBCE (up) & 4.03 & 3.96 & 3.50 & 3.77 & 3.82 & 3.83 & 3.90 & 3.93 & 3.92 & 3.92 & 3.91 \\ 
  SSCE & -- & -- & -- & -- & -- & 4.24 & 4.11 & 3.99 & 4.00 & 4.00 & 3.99 \\ 
  BSSCE & -- & -- & -- & -- & -- & 4.02 & 4.01 & 4.01 & 4.01 & 4.01 & 4.01 \\ 
  BSSL & -0.23 & 4.07 & 6.80 & 4.05 & 4.00 & 4.00 & 4.01 & 3.98 & 3.99 & 3.99 & 3.99 \\ 
   \hline
\end{tabular}\subcaption*{Second scenario: $|\gamma_j|>0$ for $j\le 10$ and $|\beta_j|>0$ for $j\le 15$}
\begin{tabular}{l|rrrrrrrrrrr}
  \hline
 & 25 & 30 & 35 & 40 & 45 & 50 & 55 & 60 & 65 & 70 & 75 \\ 
  \hline
RBCE (low) & 2.79 & 3.70 & 3.77 & 3.56 & 3.69 & 3.70 & 3.81 & 3.78 & 3.81 & 3.82 & 3.85 \\ 
  RBCE (up) & 3.65 & 4.01 & 3.96 & 3.82 & 3.90 & 3.86 & 3.92 & 3.89 & 3.91 & 3.88 & 3.91 \\ 
  SSCE & -- & -- & -- & -- & -- & 4.80 & 4.05 & 4.06 & 6.02 & 4.04 & 4.04 \\ 
  BSSCE & -- & -- & -- & -- & -- & 10.34 & 8.12 & 4.17 & 4.04 & 4.06 & 4.05 \\ 
  BSSL & -6.68 & 3.62 & 4.06 & 4.07 & 4.06 & 4.02 & 4.05 & 4.07 & 4.03 & 4.05 & 4.04 \\ 
   \hline
\end{tabular}
\end{table}

\begin{figure}
    \centering
    \includegraphics[width = 0.9\linewidth]{RBCE_comp.pdf}
    \caption{Comparison of different methods in estimating the treatment effect.}
    \label{fig:comp:trt}
\end{figure}

For the variable identification, we use the notion of
different losses as described earlier. We consider
$\ell_1=\ell_2 = 1$ and $\ell_3= 0.2$. This is a
simplified way of choosing the loss function, we can choose more 
sophisticated loss functions based on \cite{ZAFFALON20121282}. We use this associated loss
to obtain the total loss, which we present in \cref{tab:misspec}.
In the table we denote the misspecification by counting the number of
false positives (FP) and false negatives (FN). For RBCE, we have an
additional column `ID' which denotes the number of variables which remain as
indeterminate.
From the table it can be seen that for the first scenario, our method
abstains from identifying some variables for 
$n <50$. Especially for $n=25$, our method identifies 26 and 23 variables as indeterminate
for the first setting and second setting respectively. However,
later on our method gives more precise results in terms of variable
selection. We also notice that BSSL tends to perform poorly
in terms of variable selection for $n=25$, this can be seen
from the treatment effect estimation as well.
Moreover, we observe that for the second setting both SSCE
and BSSCE underperform in identifying the active variables, which can 
be explained from \cref{tab:causal} as well.

% latex table generated in R 4.3.0 by xtable 1.8-4 package
% Wed May 24 00:50:53 2023
\begin{table}[ht]
\caption{Loss based on misspecification of active variables in different
models.}\label{tab:misspec}
\centering
\subcaption*{First scenario: $|\gamma_j|, |\beta_j|>0$ for $j\le 10$}
\begin{tabular}{|c||rrr|r||rr|r||rr|r||rr|r|}
  \hline
  &\multicolumn{4}{c||}{RBCE}&\multicolumn{3}{c||}{SSCE}
  &\multicolumn{3}{c||}{BSSCE}&\multicolumn{3}{c|}{BSSL}\\
  \cline{2-14}
 Samples & FP & FN & ID & Tot & FP & FN & Tot & FP & FN & Tot & FP & FN & Tot \\ 
  \hline
25 &   0 &   0 &  26 & 5.2 & -- & -- & -- & -- & -- & -- &  12 &   2 & 14\\ 
  30 &   0 &   1 &   1 & 1.2 & -- & -- & -- & -- & -- & -- &   0 &   0 & 0\\ 
  35 &   0 &   0 &   1 & 0.2 & -- & -- & -- & -- & -- & -- &   0 &   0 & 0\\ 
  40 &   0 &   1 &   0 & 1.0 &-- & -- & -- & -- & -- & -- &   0 &   0 & 0\\ 
  45 &   0 &   0 &   1 & 0.2 & -- & -- & -- & -- & -- & -- &   0 &   0 & 0\\ 
  50 &   0 &   0 &   0 & 0.0 &  0 &   0  & 0 &   0 &   0  & 0 &   0 &   0  & 0\\ 
  55 &   0 &   0 &   0 & 0.0 &   0 &   0 &   0  & 0 &   0  & 0 &   0 &   0  & 0\\ 
  60 &   0 &   0 &   0 & 0.0 &   0 &   0 &   0 &   0  & 0  & 0 &   0 &   0  & 0\\ 
  65 &   0 &   0 &   0 & 0.0 &   0 &   0 &   0 &   0 &   0  & 0  & 0 &   0  & 0\\ 
  70 &   0 &   0 &   0 & 0.0 &   0 &   0 &   0 &   0 &   0  & 0  & 0 &   0  & 0\\ 
  75 &   0 &   0 &   0 & 0.0 &   0 &   0 &   0 &   0 &   0  & 0  & 0 &   0  & 0\\ 
   \hline
\end{tabular}
\subcaption*{Second scenario: $|\gamma_j|>0$ for $j\le 10$ and $|\beta_j|>0$ for $j\le 15$}
\begin{tabular}{|c||rrr|r||rr|r||rr|r||rr|r|}
  \hline
  &\multicolumn{4}{c||}{RBCE}&\multicolumn{3}{c||}{SSCE}
  &\multicolumn{3}{c||}{BSSCE}&\multicolumn{3}{c|}{BSSL}\\
  \cline{2-14}
 Samples & FP & FN & ID & Tot & FP & FN & Tot & FP & FN & Tot & FP & FN & Tot \\ 
  \hline
25 &   0 &   2 &  23 & 6.6 & -- & -- & -- & -- & -- & -- &   9 &   4 & 13\\ 
  30 &   0 &   2 &   9 & 3.8 & -- & -- & -- & -- & -- & -- &   0 &   0 & 0 \\ 
  35 &   0 &   0 &  18 & 3.6 & -- & -- & -- & -- & -- & -- &   0 &   0 & 0 \\ 
  40 &   0 &   0 &   5 & 1.0 &-- & -- & -- & -- & -- & -- &   0 &   0 & 0 \\ 
  45 &   0 &   0 &   3 & 0.6 & -- & -- & -- & -- & -- & -- &   0 &   0 & 0 \\ 
  50 &   0 &   0 &   1 & 0.2 &  0 &   7 & 7 &   0 &  14 & 14 &   0 &   0 & 0 \\ 
  55 &   0 &   0 &   1 & 0.2 &  0 &   0 & 0 &  0 &  12 & 12 &  0 &   0 & 0 \\ 
  60 &   0 &   0 &   1 & 0.2 &  1 &   0 & 1 &  0 &   0 &  0 & 0 &  0 &   0 \\ 
  65 &   0 &   0 &   0 & 0.0 &  0 &  12 & 12 &  0 &   0 &  0 & 0 &  0 &   0 \\ 
  70 &   0 &   0 &   0 & 0.0 &  0 &   0 &  0 &  0 &   0 &  0 & 0 &  0 &   0 \\ 
  75 &   0 &   0 &   0 & 0.0 &  0 &   0 &  0 &  0 &   0 &  0 & 0 &  0 &   0 \\ 
   \hline
\end{tabular}
\end{table}

\section{Conclusion}\label{sec:conc}

Causal effect estimation is an important tool in statistical learning and needs to
be performed with utmost care as in many cases we may have severe consequence of poor estimation.
In this paper, we tackle this issue by proposing a robust Bayesian analysis of causal 
effect estimation problem for high dimensional data. Our 
framework is focused on the effect of prior elicitation on confounder selection
as well as causal effect estimation. We consider a spike and slab type
prior for confounder selection and discuss the possible sources of uncertainty that
need to be tackled carefully. We were particularly focused on the uncertainty associated
with prior selection probabilities for which we consider a set of beta priors to perform
sensitivity analysis. We showed that the sensitivity analysis on the prior selection probability
gives us a robust confounder selection scheme. In this way, we can abstain from selecting
a confounder when the available data is not sufficient. We also propose a generalised
utility based framework, where we associate a loss for abstaining which can be interpreted 
as the cost of further data collection. Finally, we illustrate our method with synthetic dataset
and compare with other state of the art Bayesian methods.

Currently, the paper proposes a robust Bayesian approach for causal effect estimation where
we rely on sampling strategies to obtain the posterior bounds as well as performing 
variable selection. In future, it will be interesting to derive inner approximation bounds
for the posterior estimates to reduce the computational cost. Moreover, for the sake of
illustration, we rely on simple loss functions and elicitation strategy. In future, we would
like to investigate different elicitation strategies for the method and explore alternative 
loss functions for formulating a decision theoretic framework. Last but not the least,
we noticed that our method
is in good agreement with other methods with an added level of robustness.
This confirms that our method has good potential for real-life problems,
and we intend to apply it on a real dataset in future work.

\bibliographystyle{splncs04}
\bibliography{basu_causal}

\end{document}